Title

### 2012-05-21 inc/ms_model/ms_model.tex
Improved generalized extension for external forces
Move Liouville's theorem implications to extension (ie my contributions) section

Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
These terms will be denoted $K^{flow}$ and $K^{int}$ respectively.
\begin{equation} \label{eq:gaussian_integral_dt_k}
  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
\end{equation}
Collecting and expanding terms where possible gives these forms of the $K$ terms
\begin{gather}
  K^{flow}_{ij} \equiv -\frac{1}{N} \int u_i u_j \frac{\vec{p}}{m} \frac{\partial f}{\partial \vec{r}} d\vec{r} d\vec{p} \\
  K^{int}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
  n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
  \begin{split}
    \Phi(\vec{r};t) \equiv & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \\
    = & \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \int d\vecprime{r} V(\vec{r} - \vecprime{r}) \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right)
  \end{split}
\end{gather}
Obviously the $K^{flow}$ term is far easier to evaluate
\begin{equation}
  K^{flow} = 
  \begin{pmatrix}
    k^{flow}_{\smallT} & 0 & 0 \\
    0 & k^{flow}_{\smallT} & 0 \\
    0 & 0 & k^{flow}_{z}
  \end{pmatrix}
  \quad \text{and} \quad
  k^{flow}_{\alpha} = 
  \begin{pmatrix}
    \dfrac{2\sigma_{\alpha}}{m} & \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) \\
    \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) & 0
  \end{pmatrix}
\end{equation}
These analyses will be referred to while deriving the influence of external forces (Section \ref{sec:external_forces}).

@@

\subsection{Implications of Liouville's Theorem}

One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
%TODO find a good reference for this

To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
Using \ref{eq:ag_original} it is easy to show that
\begin{gather}
  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
  \label{eq:define_big_gamma}
  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
\end{gather}
This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.

Not only is this an interesting result, but it can be both instructive and computationally useful.
By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substituting into \ref{eq:ag_original}.
\begin{subequations} \label{eq:ag_big_gamma}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
\end{gather}
\end{subequations}
these four equations now contain all of the information previously contained in six. Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
This will become more apparent when in future sections
%TODO include internal reference
it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

>>>>

Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
These terms will be denoted $K^{flow}$ and $K^{force}$ respectively.
\begin{equation} \label{eq:gaussian_integral_dt_k}
  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
\end{equation}
Collecting and expanding terms where possible gives these forms of the $K$ terms
\begin{gather}
  K^{flow}_{ij} \equiv -\frac{1}{N} \int u_i u_j \frac{\vec{p}}{m} \frac{\partial f}{\partial \vec{r}} d\vec{r} d\vec{p} \\
  \label{eq:Kforce}
  K^{force}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
  n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
    \label{eq:potential_integral}
    \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t)
\end{gather}
Obviously the $K^{flow}$ term is far easier to evaluate
\begin{equation}
  K^{flow} = 
  \begin{pmatrix}
    k^{flow}_{\smallT} & 0 & 0 \\
    0 & k^{flow}_{\smallT} & 0 \\
    0 & 0 & k^{flow}_{z}
  \end{pmatrix}
  \quad \text{and} \quad
  k^{flow}_{\alpha} = 
  \begin{pmatrix}
    \dfrac{2\sigma_{\alpha}}{m} & \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) \\
    \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) & 0
  \end{pmatrix}
\end{equation}

@@
